### 2023-09-06 ms/ms.qmd
initial commit all files

---
title: Amphistomy increases leaf photosynthesis more in coastal than montane plants of Hawaiian ʻilima (*Sida fallax*)
format:
  pdf:
    keep-tex: true  
    include-in-header: 
      - preamble.tex
    toc: false
header-includes: |
  \addtokomafont{disposition}{\rmfamily}
  \RedeclareSectionCommand[
    font=\normalfont\Large]{section}
  \RedeclareSectionCommand[
    font=\normalfont\normalsize\bfseries]{subsection}
  \RedeclareSectionCommand[
    font=\normalfont\normalsize\itshape]{subsubsection}
  \RedeclareSectionCommand[
    font=\normalfont\normalsize]{paragraph}
mainfont: Times New Roman
bibliography: stomata-ilima.bib
csl: https://www.zotero.org/styles/american-journal-of-botany
---

\begin{center}
Genevieve Triplett$^1$, Thomas N. Buckley$^2$, Christopher D. Muir$^{1,3,*}$
\end{center}

$^1$ School of Life Sciences, University of Hawaiʻi Mānoa, Honolulu, HI, USA 96822

$^2$ School of Life Sciences, University of California, Davis, CA, USA 95616

$^3$ Department of Botany, University of Wisconsin, Madison, WI, USA 53706

$^*$ Corresponding author: Christopher D. Muir, cdmuir@wisc.edu

Manuscript received _______; revision accepted _______.

Running head: Amphistomy advantage in ʻilima

\newpage

```{=html}

<!-- TO DO,

1. at end of INTRODUCTION, address Tom's comments

"suggest you rigorously define sun leaves and shade leaves here, and dispense with the quotes."

Brief Communications—Brief Communications are short (c 3000–4000 word) research articles reporting exciting, significant new findings. They include no more than 4 figures and tables, combined. Manuscripts, and their abstracts, should be organized as described for Research articles. Authors must adhere to the same membership and page charge requirements as for regular Research Articles.

a. Title and Running Heads
The manuscript title for research papers should be specific and informative, conveying the key findings of the research in an active voice. Provide a title written with sentence-style capitalization. Below the title, list authors: each author’s first name, middle initial, surname. On the next line, give affiliation and unabbreviated address. If authors have different affiliations and addresses, add a superscript number after each author’s name to indicate the footnoted address. Include another footnote superscript number to indicate the author for correspondence.

Include the line: Manuscript received _______; revision accepted _______.

After this text, add a short title no longer than 65 characters (including spaces) to be used as a running head in the article.

b. Abstract
AJB requires structured abstracts for Research Articles and Brief Communications. The abstract is 250 words or fewer, written in the following structured format:
• Premise of the study (why the work was done, what major questions of plant biology are addressed, and why it is important to the broad AJB readership)
• Methods
• Key results
• Conclusions (what major points should the reader take from this article)

The abstract may be all a reader sees initially, so it should be written to capture the interest of the general botanical community as well as specialists. Avoid references; if essential, cite parenthetically with journal name, volume number, pages, and year.
-->
```
```{r, echo = FALSE, message = FALSE, warning = FALSE}

library(brms)
library(dplyr)
library(ggdist)
library(glue)
library(readr)
library(stringr)
library(tidyr)

licor = read_rds("../processed-data/licor.rds")

# Function to print estimates and CIs
pretty_est = function(estimate, lower, upper, ...) {
  .e = prettyNum(estimate, ...)
  .l = prettyNum(lower, ...)
  .u = prettyNum(upper, ...)
  # glue("{.e}~[{.l}-{.u}]")
  glue("{.e}~[\\numrange{{{.l}}}{{{.u}}}]")
}

# From 04_fit-licor.R
fit_licor = read_rds("../objects/fit_licor.rds")
dim_fit_licor = dim(as.array(fit_licor$fit))

# From 05_fit-traits.R
fit_stomata = read_rds("../objects/fit_stomata.rds")
fit_thickness = read_rds("../objects/fit_thickness.rds")

dim_fit_stomata = dim(as.array(fit_stomata$fit))
dim_fit_thickness = dim(as.array(fit_thickness$fit))

habitat_thickness = fit_thickness |>
  as_draws_df(c("b_Intercept", "b_site_typemontane")) |>
  transmute(
    coastal = exp(b_Intercept),
    montane = exp(b_Intercept + b_site_typemontane),
    diff = coastal - montane
  ) |>
  point_interval(.width = 0.95, .point = median, .interval = qi)

thickness_diff = pretty_est(habitat_thickness$diff,
                     habitat_thickness$diff.lower,
                     habitat_thickness$diff.upper,
                     digits = 2L)

# From 08_plot-habitat-aa.R
# Estimated difference in AA between habitats
habitat_aa = read_rds("../objects/habitat_aa.rds") |>
  pivot_wider(names_from = site_type, values_from = AA) |>
  mutate(AA_diff = coastal - montane, times_increase = coastal / montane) |>
  point_interval(coastal, montane, AA_diff, times_increase,
                 .width = 0.95, .point = median, .interval = qi)

AA_coastal = pretty_est(habitat_aa$coastal,
                        habitat_aa$coastal.lower,
                        habitat_aa$coastal.upper,
                        digits = 2L)

AA_montane = pretty_est(habitat_aa$montane,
                        habitat_aa$montane.lower,
                        habitat_aa$montane.upper,
                        digits = 2L)

AA_diff = pretty_est(habitat_aa$AA_diff,
                     habitat_aa$AA_diff.lower,
                     habitat_aa$AA_diff.upper,
                     digits = 2L)

habitat_aa1 = read_rds("../objects/habitat_aa1.rds") |>
  pivot_wider(names_from = site_type, values_from = AA) |>
  mutate(AA_diff = coastal - montane, times_increase = coastal / montane) |>
  point_interval(coastal, montane, AA_diff, times_increase,
                 .width = 0.95, .point = median, .interval = qi)

AA_diff1 = pretty_est(habitat_aa1$AA_diff,
                     habitat_aa1$AA_diff.lower,
                     habitat_aa1$AA_diff.upper,
                     digits = 2L)

# From 09_plot-thickness-aa.R
coef_thickness_aa = read_rds("../objects/coef_thickness_aa.rds") |>
  point_interval(slope, .width = 0.95, .point = median, .interval = qi)

thickness_aa = pretty_est(
  coef_thickness_aa$slope,
  coef_thickness_aa$.lower,
  coef_thickness_aa$.upper,
  digits = 2L
)

# From 10_calculate-gmaxratio.R
gmaxratio_sitetype = read_rds("../objects/gmaxratio_sitetype.rds") |>
  select(-lower_gmax, -upper_gmax) |>
  pivot_wider(names_from = "habitat", values_from = "gmax_ratio") |>
  mutate(d = coastal - montane) |>
  point_interval(d, .width = 0.95, .point = median, .interval = qi) |>
  with(pretty_est(d, .lower, .upper, digits = 2L))

# From 12_plot-gmaxratio-aa.R
coef_gmaxratio_aa = read_rds("../objects/coef_gmaxratio_aa.rds") |>
  point_interval(slope, .width = 0.95, .point = median, .interval = qi)

gmaxratio_aa = pretty_est(
  coef_gmaxratio_aa$slope,
  coef_gmaxratio_aa$.lower,
  coef_gmaxratio_aa$.upper,
  digits = 2L
)

# From 13_plot-habitat-Ags.R
fit_habitat_Ags = read_rds("../objects/fit_habitat_Ags.rds")

# Calculation amount of gsw extrapolation
gsw_extrap = licor |>
  summarize(min = min(gsw), max = max(gsw), .by = c("site_code", "aperture")) |>
  pivot_wider(names_from = aperture, values_from = min:max) |>
  rowwise() |>
  mutate(gsw = ifelse(
    `max_two-sided` > `max_one-sided`,
    mean(c(`max_one-sided`, `min_two-sided`)),
    mean(c(`min_one-sided`, `max_two-sided`))
  )) |>
  transmute(site_code = site_code, d_gsw = (`min_two-sided` - `max_one-sided`) / 2) |>
  filter(d_gsw > 0) |>
  pull(d_gsw) %>%
  sprintf("%.2e", .) |>
  str_replace("^([0-9].[0-9]{2})e-0([0-9])", "\\1\\\\times 10^{-\\2}") |>
  str_c(collapse = "~\\text{and}~") %>%
  paste0("$", ., "~\\text{mol}~\\text{m}^{-2}~\\text{s}^{-1}$")

```

# ABSTRACT --- {#abstract}

\noindent \textbf{Premise of the study}  The adaptive significance of stomata on both upper and lower leaf surfaces, called amphistomy, is unresolved. A widespread association between amphistomy and open, sunny habitats suggests the adaptive benefit of amphistomy may be greatest in these contexts, but this hypothesis has not been tested experimentally. Understanding why amphistomy evolves can inform  its potential as a target for crop improvement and paleoenvironment reconstruction.

\textbf{Methods}  We developed a new method to quantify "amphistomy advantage", $\mathrm{AA}$, as the ratio of photosynthesis in an amphistomatous leaf to that of the same leaf with the same total stomatal conductance but with gas exchange blocked through the upper (adaxial) surface, which we term ʻpseudohypostomyʻ. We thus estimated $\mathrm{AA}$ and related physiological and anatomical traits in 12 populations, six coastal (open, sunny) and six montane (closed, shaded), of the indigenous Hawaiian species ʻilima (*Sida fallax*).

\textbf{Key results}  Coastal ʻilima leaves benefit 4.58 times more from amphistomy compared to their montane counterparts. Our evidence was equivocal with respect to two hypotheses – that coastal leaves benefit more because they are thicker and therefore have lower CO$_2$ conductance through the internal airspace, and that leaves with more similar conductances at each surface benefit more than leaves with few stomata on the upper surface.

\textbf{Conclusions}  This is the first direct experimental evidence that amphistomy *per se* increases photosynthesis, consistent with the hypothesis that parallel pathways through upper and lower mesophyll increase the supply of CO$_2$ to chloroplasts. The prevalence of amphistomatous leaves in open, sunny habitats can partially be explained the increased benefit of amphistomy in ʻsunʻ leaves, but the mechanistic basis of this observation is an area for future research.

**Keywords:** amphstomy, leaf, light, Hawaiʻi, leaf, *Sida fallax*, stomata

# INTRODUCTION --- {#introduction}

Amphistomy, the presence of stomata on both lower and upper surfaces of broad leaves, should increase carbon gain by reducing the average diffusion pathlength between stomata and chloroplasts (\autoref{fig:compare-circuit-diagrams}), yet paradoxically this seemingly simple adaptation is uncommon in nature and we don't know why. Understanding variation in stomatal traits like amphistomy is imperative because these tiny pores play an outsized ecological role in the global carbon and water cycles [@hetherington_role_2003; @berry_stomata:_2010]. A widely applicable, accurate representation of how stomata mediate the relationship between CO$_2$ gained through photosynthesis and water lost through evapotranspiration is essential to predict future climate using Earth Systems Models [@jarvis_interpretation_1976; @biggins_model_1987; @collatz_physiological_1991; @leuning_critical_1995; @sellers_modeling_1997]. Optimality models accurately predict the major cause of water loss, stomatal conductance ($g_\mathrm{sw}$), by assuming plants maximize carbon gain minus a cost of water [@cowan_stomatal_1977; @givnish_economy_1986; @medlyn_reconciling_2011; @lin_optimal_2015; @wang_towards_2017; @franks_comparing_2018; @deans_optimization_2020; @franklin_organizing_2020; @wang_theoretical_2020; @harrison_ecoevolutionary_2021]. Despite the success of optimality modeling in predicting $g_\mathrm{sw}$, the same modeling approach has so far failed to explain the rarity of amphistomatous leaves [@muir_is_2019]. **This gap between theory and observations strongly implies that we remain ignorant about some key benefits and costs associated with stomata.**

Where are amphistomatous leaves found and why aren't they more common? Among terrestrial flowering plants, amphistomatous leaves are rarely found on woody plants and shade-tolerant herbs, but they are common in annual and perennial herbs from sunny habitats [@salisbury_i_1928; @parkhurst_adaptive_1978; @mott_adaptive_1982; @peat_comparative_1994; @gibson_structure-function_1996; @jordan_using_2014; @muir_making_2015; @bucher_stomatal_2017; @muir_light_2018]. Even in resupinate leaves where the abaxial surface faces up toward the sky, stomata develop on the lower adaxial surface [@lyshede_comparative_2002]. Exceptions to this general pattern include some arid woody plants which typically have vertically oriented, isobilateral leaves [@wood_physiology_1934; @jordan_using_2014; @de_boer_apparent_2016; @drake_two_2019] and floating/amphibious leaves of aquatic plants [@kaul_anatomical_1976; @doll_callitriche_2021]. The dearth of amphistomatous leaves should be quite surprising and has been described as one of the most important unsolved problems in the study of leaf structure-function relations despite some recent progress [@grubb_leaf_1977; @grubb_leaf_2020].

Amphistomatous leaves should be common because, all else being equal, a leaf with a given number of stomata per area could increase its photosynthetic rate simply by apportioning approximately half its stomata to each surface [@parkhurst_adaptive_1978; @gutschick_photosynthesis_1984; @gutschick_photosynthesis_1984-1]. The key difference between a hypo- and amphistomatous leaf, holding all other factors constant, is that an amphistomatous leaf has two parallel diffusion paths through the internal airspace to any given chloroplast. Those airspaces pose a resistance for CO$_2$ diffusion, so CO$_2$ concentration drops as it approaches chloroplasts. Shorter pathways mean a smaller drop in CO$_2$ concentration. Thus, chloroplasts in amphistomatous leaves experience higher CO$_2$ concentrations than in hypostomatous leaves, thereby increasing photosynthesis. The airspace resistance (or its inverse, the airspace conductance, $g_\mathrm{ias}$) is rarely measured directly and there is disagreement between empirical [@parkhurst_intercellular_1990; @morison_lateral_2005; @evans_resistances_2009; @tomas_importance_2013; @earles_beyond_2018; @santrucek_partitioning_2019; @nobel_physicochemical_2020; @harwood_understanding_2021; @marquez_assessing_2023] and theoretical models [@tholen_mechanistic_2011; @ho_three-dimensional_2016; @theroux-rancourt_maximum_2021]. The $g_\mathrm{ias}$ in thin, porous leaves may be so large as to be inconsequential given much lower conductances for other components of the diffusion pathway, whereas the $g_\mathrm{ias}$ of thick leaves with little airspace may greatly hinder CO$_2$ diffusion to chloroplasts. Amphistomy should confer the largest photosynthetic benefit in leaves with intrinsically low $g_\mathrm{ias}$.

Amphistomy should also enhance photosynthesis when leaf boundary layer resistance is high, because apportioning total flux between two boundary layers rather than one results in a smaller CO$_2$ concentration drop between the atmosphere and stomata. A similar effect has been validated with a computer model and measurements for transpiration: amphistomatous leaves lose somewhat more water for the same vapor pressure deficit and total $g_\mathrm{sw}$ [@foster_influence_1986], but the additional carbon gain should be enough to offset this cost under most realistic conditions [@muir_is_2019].

The most promising adaptive hypothesis is that amphistomy is important for maximizing photosynthetic rate under high light. @mott_adaptive_1982 proposed that "plants with a high photosynthetic capacity, living in full-sun environments, and experiencing rapidly fluctuating or continuously available soil water" would benefit most, in terms of increased carbon gain, from having amphistomatous leaves. As described above, herbs from sunny habitats are often amphistomatous. Most variation in stomatal density ratio ($\mathrm{SR}$, the ratio of stomatal density between the upper and lower surfaces) among species is assumed to be genetic, but there is also putatively adaptive plasticity in response to light. Leaves of *Ambrosia cordifolia*, a desert perennial herb, are hypostomatous under low light (photosynthetic photon flux density, PPFD = 110 $\mu \text{mol}~\text{m}^{-2}~\text{s}^{-1}$) but develop \~20% of their stomata on the upper surface under high light (1700 $\mu \text{mol}~\text{m}^{-2}~\text{s}^{-1}$) [@mott_amphistomy_1991]. Similarly, *Solanum lycopersicum* leaves are hypostomatous when grown in the shade but develop \~20% of their stomata on the upper surface grown under high light-intensity [@gay_influence_1975]. Adult leaves of *Eucalyptus globulus* are amphistomatous, but the proportion of adaxial stomata increases from \~10-20% under low light to \~30-40% under high light [@james_leaf_2001]. In summary, both genetic and plastic responses to evince a widespread association between light and $\mathrm{SR}$.

The association between high light and amphistomy suggests that 'sun' leaves have the most to gain in terms of increased photosynthesis from having stomata on both surfaces, as @mott_adaptive_1982 hypothesized. @parkhurst_adaptive_1978 proposed quantifying this benefit as 'amphistomy advantage' ($\mathrm{AA}$), which we adopt here with some modification (see Materials and Methods). This hypothesis has never been tested directly by comparing the photosynthetic rate of an amphistomatous leaf to that of an otherwise identical hypostomatous leaf with the same total stomatal conductance under the same conditions. We propose a straightforward method to do this by experimentally creating a pseudohypostomatous leaf with gas exchange blocked through the upper surface (see Materials and Methods). One reason that sun leaves might have greater $\mathrm{AA}$ is that they are usually thicker or denser [@poorter_metaanalysis_2019], resulting in lower $g_\mathrm{ias}$ [@parkhurst_adaptive_1978]. A nonmutually exclusive hypothesis is that if sun leaves have a stomatal density ratio closer to 0.5 (same density on each leaf surface), this will confer a greater advantage than an amphistomatous leaf with most stomata on one surface. In other words, amphistomy doesn't make much difference if one leaf surface has very open stomata on it. We therefore predict that sun leaves will have greater $\mathrm{AA}$ possibly because they have thicker leaves and/or $\mathrm{SR}$ closer to 0.5. We actually report $g_{\mathrm{smax,ratio}}$, which is similar to $\mathrm{SR}$ except that takes into account differences in both stomatal density and size between surfaces.

The native flora of the Hawaiian archipelago is a excellent system to test the relationship between light habitat and $\mathrm{AA}$. Many lineages have adapted to different light habitats after colonization and leaf anatomical traits such as $\mathrm{SR}$ and thickness vary within and among closely related species. It is hypothesized that the common ancestor in many Hawaiian clades was a "weedy" species with high dispersal ability adapted to open habitats [@carlquist_biota_1966]. Colonization was followed by adaptive radiation into higher elevation, montane, closed, forested habitats. Consequently, adaptation to sun and shade is a common axis of phenotypic variation among Hawaiian plants such as lobeliads [@givnish_adaptive_2004; @montgomery_adaptive_2008; @givnish_origin_2009; @givnish_common-garden_2014; @scoffoni_light-induced_2015], *Bidens* [@carlquist_biota_1966; @knope_dispersal_2020], *Scaevola* [@robichaux_evolution_1984; @mckown_trait_2016], *Euphorbia* [@sporck_hawaiian_2011], and *Plantago* [@dunbar-co_leaf_2009].

Here we focus on variation within an indigenous plant species *Sida fallax* Walp. (Malvaceae), known in the Hawaiian language as ʻilima. ʻIlima is found from sea level to elevations $>1000$ mas on multiple Hawaiian islands. Coastal populations are morphologically different from montane populations (\autoref{fig:study-system}). Coastal regions of Hawaiʻi are characterized by high sun exposure, warmer temperatures, high winds, salinity, and variation in water availability. Coastal populations of ʻilima tend to be short and prostrate which likely helps them to withstand the windy environment (\autoref{fig:study-system}a). The leaves of these populations are covered on both surfaces in dense, soft hairs that give the leaves a silvery green appearance (\autoref{fig:study-system}b), which helps mitigate water loss by reflecting solar radiation, thereby lowering leaf temperature [@ehleringer_pubescence_1978]. Montane regions, on the other hand, provide very different challenges. Many other tall species grow on the slopes of these wet mountainous regions, which makes light competition a factor that plants may need to adapt to. Possibly due to this, montane populations are erect and shrub- or tree-like, capable of growing meters tall with strong, woody stems. These individuals have smooth, green foliage with serrated edges. Montane populations exhibit traits that help them to compete for light availability (Raunch et al. 1997). This montane morphology is not found in *S. fallax* populations on other Pacific Islands [@pejhanmehr_phylogenetic_2022]. The variable morphology between coastal and montane populations has even led some scientists to argue that it makes up a species complex, rather than a single species (Raunch et al. 1997).

The objectives of our study are to test whether 1) 'sun' leaves of coastal ʻilima plants will have greater $\mathrm{AA}$ than 'shade' leaves of montane plants; and if so, is this because 2a) coastal plants have thicker leaves than montane plants and/or 2b) coastal plants have a $g_{\mathrm{smax,ratio}}$ closer to 0.5?

```{=html}
<!-- TEXT FROM PROPOSAL THAT MAY HAVE SOMETHING USEFUL:

*Knowing the adaptive significance of amphistomy will advance our understanding of leaf structure-function relationships, aid reconstructing paleoenvironments, and guide crop optimization*. Stomata are studied in disparate fields such as cell biology, ecosystem ecology, and paleontology [@hetherington_role_2003; @bergmann_stomatal_2007; @berry_stomata:_2010; @mcelwain_paleoecology_2017]. The intense focus on stomata makes it all the more surprising that we know relatively little about the adaptive significance of where stomata are on the leaf! Amphistomy is deeply integrated with leaf traits associated with high photosynthetic capacity and leaf water transport capacity [@smith_leaf_1997; @drake_two_2019; @xiong_one_2020]. Knowing the upper limits of photosynthesis will help breed more productive crops [@long_can_2006] and predict future terrestrial carbon sinks at higher atmospheric CO$_2$ [@keenan_terrestrial_2018]. In paleontology, stomata visible on fossil leaves help reconstruct paleo-CO$_2$ concentrations based on density/index [@konrad_modelling_2008; @franks_co2-forced_2009; @mcelwain_paleoecology_2017; but see @jordan_links_2020], but stomatal density ratio ($\mathrm{SR}$) may add information about whether plants lived in open or closed vegetation [@salisbury_i_1928; @jordan_using_2014; @bucher_stomatal_2017]. Finally, $\mathrm{SR}$ tends to increase (higher proportion of stomata on upper surface) during domestication [@milla_shifts_2013], but we do not know why. Our work will help infer how changes in $\mathrm{SR}$ and other leaf anatomical traits affect photosynthesis and water use in crops. This is particularly promising because we may learn ways to increase leaf photosynthesis without increasing water loss by altering $\mathrm{SR}$ and/or the regulation of stomatal aperture on each leaf surface.

To redress these gaps in our knowledge, in Aim 1 we will model the benefits and costs of amphistomy under low- and high-light conditions and in Aims 2-3 we will test model predictions using the wild tomatoes as model clade (*Solanum* spp.; Solanaceae). Specifically, we will experimentally manipulate whether gas exchange proceeds through one or both leaf surfaces, holding all else constant. This will enable us to measure how much amphistomy increases photosynthesis compared to an otherwise identical hypostomatous leaf (Aim 2). In Aim 3, we will test whether amphistomy is costly in shade leaves because it causes dehydration of the palisade mesophyll under high evaporative demand. The result of this project will determine whether high light, photosynthetic capacity, and adequate hydration are sufficient to explain the distribution of amphistomatous leaves in nature, as Mott and colleagues hypothesized 40 years ago.

The dearth of amphistomatous leaves implies a fitness cost associated with stomata on the upper leaf surface in particular [@parkhurst_adaptive_1978; @gutschick_photosynthesis_1984-1; @muir_is_2019]. Two nonmutually exclusive hypotheses for the cost of upper stomata have received the most attention, but neither has been tested experimentally. We will test one hypothesis, which is that amphistomy hinders photosynthesis by causing dehydration of the palisade mesophyll under high evaporative demand [@buckley_sites_2017; @drake_two_2019]. This cost of amphistomy could be greater in shade leaves because they dehydrate severely when exposed to intense sunflecks, warm temperatures, and/or low humidity, whereas sun leaves have constitutive traits that mitigate the risk of dehydration. Sensitivity to dehydration would be more consequential if it affected the palisade mesophyll, which is adjacent to the illuminated upper surface and is thus more photosynthetically active [@smith_leaf_1997; @evans_profiles_2003; @earles_embracing_2019] than the spongy mesophyll, which is adjacent to the lower surface (\autoref{fig:compare-circuit-diagrams}). An alternative hypothesis with some empirical support is that stomata on the upper surface are more likely to be colonized by bacterial or fungal pathogens, presumably because a propagule is more likely to land on the upper surface [@gutschick_photosynthesis_1984-1; @mckown_association_2014; @mckown_role_2019; @fetter_growthdefense_2021]. The current proposal does not address this pathogen susceptibility hypothesis because PI Muir is seeking funding for this work elsewhere. Although not the focus of our aims, our modeling and data can address two less well theorized alternative hypotheses that upper stomata compete for space on the epidermis [@de_boer_optimal_2016; @haworth_allocation_2018; @muir_how_2023] or alter palisade mesophyll structure [@dow_disruption_2017; @lundgren_mesophyll_2019; @baillie_developmental_2020].

*Despite the fact that we do not know the adaptive significance of amphistomy, the distribution of hypo- and amphistomatous leaves is nonrandom and most likely caused by natural selection.* Consistent relationships between stomatal traits, growth form, and light environment strongly suggest convergent evolution, a hallmark of adaptive evolution. Hypo- and amphistomatous leaves are two ends of a continuum in $\mathrm{SR}$. Hypostomatous leaves have a $\mathrm{SR}$ close to 0; fully amphistomatous leaves with equal densities on each leaf have a $\mathrm{SR}$ close to 0.5. Herbaceous plants from high-light habitats tend to be amphistomatous and have higher $\mathrm{SR}$ values than herbaceous plants from shaded habitats or woody species [@salisbury_i_1928; @parkhurst_adaptive_1978; @mott_adaptive_1982; @parkhurst_intercellular_1990; @peat_comparative_1994; @muir_making_2015; @bucher_stomatal_2017; @muir_light_2018]. A second line of evidence for adaptation is that the distribution of $\mathrm{SR}$ among species is inconsistent with the expected distribution for a ratio trait evolving neutrally; instead, a nonneutral bimodal distribution evolves repeatedly in land plants [@muir_making_2015]. Finally, amphistomatous leaves are strongly coordinated with other traits involved with high photosynthetic capacity [@beerling_evolutionary_1996; @xiong_one_2020] and domestication [@milla_shifts_2013], suggesting that amphistomy is favored by selection in some circumstances. What specific fitness costs and benefits underlie these adaptive patterns?

*The most promising adaptive hypothesis is that amphistomy is important for maximizing photosynthetic rate under high light*. To reiterate, herbaceous leaves in high light habitats tend to be amphistomatous. Most variation in $\mathrm{SR}$ among species is assumed to be genetic, but there is also putatively adaptive plasticity in response to light. Leaves of *Ambrosia cordifolia*, a desert perennial herb, are hypostomatous under low light (photosynthetic photon flux density, PPFD = 110 $\mu \text{mol}~\text{m}^{-2}~\text{s}^{-1}$) but develop ~20% of their stomata on the upper surface under high light (1700 $\mu \text{mol}~\text{m}^{-2}~\text{s}^{-1}$) [@mott_amphistomy_1991]. Similarly, *Solanum lycopersicum* leaves are hypostomatous when grown in the shade but develop ~20% of their stomata on the upper surface grown under high light-intensity [@gay_influence_1975]. Adult leaves of *Eucalyptus globulus* are amphistomatous, but the proportion of upper stomata increases from ~10-20% under low light to ~30-40% under high light [@james_leaf_2001]. In summary, both genetic and plastic responses to evince a widespread association between light and $\mathrm{SR}$.

*We don’t know whether amphistomy is an adaptation to high light intensity because no one has shown experimentally that amphistomy* per se *increases CO$_2$ assimilation more under high light than under low light*. Other factors correlated with light habitat could drive selection on $\mathrm{SR}$ and result in a spurious association between light habitat and $\mathrm{SR}$. Even if light intensity is the primary causal selective agent for $\mathrm{SR}$, its effect on selection for higher/lower $\mathrm{SR}$ could be the result of 1) *facultative* responses to light intensity and/or 2) *constitutive* differences between shade and sun leaves (\autoref{fig:causal-graph}). The *facultative* hypothesis is that amphistomy increases photosynthesis under high light intensity because of high CO$_2$ demand and/or CO$_2$ supply is constrained by low $g_\text{ias}$. Under this hypothesis, the advantage of amphistomy depends on the immediate light intensity, not the light environment in which plants develop. The nonmutually exclusive *constitutive* hypothesis is that developmental plasticity results in 'sun' and 'shade' leaves with manifold differences in anatomy and biochemistry such as thicker leaves, denser cell packing, or increased Rubisco concentration [@terashima_why_2001; @lambers_plant_2008]. Sun leaves benefit more from amphistomy than shade leaves under all light intensities because they have lower $g_\text{ias}$ and/or greater CO$_2$ demand. The proposed research will quantify the relative contributions of facultative and constitutive effects on the photosynthetic advantage of amphistomy. We will describe predictions of these hypotheses in more detail below.

MOVE THIS TO INTRO Parkhurst (1978) introduced the concept of “amphistomy advantage” to quantify the photosynthetic advantage of an amphistomatous leaf over an otherwise identical hypostomatous leaf. An important limitation of his approach is that he assumed photosynthesis was simply proportional to the total diffusive conductance from the atmosphere to the sites of carboxylation; that tacitly assumes CO$_2$ concentration at the sites of carboxylation is zero, which in turn infinite local did not model assumes CO$_2$ demand is locally infinite throughout the leaf. In reality, photosynthesis at any given point in the leaf depends on the local concentrations of CO$_2$, chlorophyll, and Rubisco, and the local irradiance. Thus, some mesophyll layers can draw down chloroplastic CO$_2$ concentration to a low level and potentially assimilate a lot of carbon when CO$_2$ is adequately supplied, whereas other layers have much weaker CO$_2$ demand. By not incorporating a CO$_2$ demand function, Parkhurst assumed that any increase in CO$_2$ supply was matched by a concomitant increase in CO$_2$ assimilation. We now know that increasing assimilation depends on both CO$_2$ supply and demand. If supply increases, but the demand function is flat because of low light intensity and/or low photosynthetic capacity, then $A$ won’t change much (demand-limited). Conversely, if the demand function is steep because there is excess light and/or enzymatic capacity, then increased CO$_2$ supply will yield a large increase in $A$ (supply-limited).

 -->
```
```{=tex}
\begin{figure}[H]
  \includegraphics{../figures/study-system.pdf}
  \caption{(A) Typical growth form of montane (left) and coastal (right) ‘ilima plants and (B) leaves. (C) Map of the sites that were sampled on the islands of Oʻahu and Hawaiʻi (aka Big Island).}
  \label{fig:study-system}
\end{figure}
```
# MATERIALS AND METHODS --- {#methods}

## Plant sampling --- {#plant-sampling}

We identified 8 suitable natural populations of ʻilima on Oʻahu and 5 on Hawaiʻi Island by consulting [@yorkston_interfertility_2006] and citizen scientist records on iNaturalist [@noauthor_inaturalist_2022] (\autoref{fig:study-system}C; \autoref{tbl-locations}). We avoided sites that appeared to be cultivated. We visited sites between August and November 2022. For logistical reasons, the sites on Hawaiʻi were sampled in one three-day trip. We haphazardly sampled eight plants distributed evenly between the highest and lowest elevation plants along a transect at each site. For safety and conservation reasons, transects were along a trail or road. We did not sample small individuals if there was risk removing leaves would cause mortality. From each plant, we collected two fully expanded leaves for traits measurements. We sampled stomatal traits on all leaves; leaf thickness on one leaf from three randomly selected plants per site; and, due to limited time, a single leaf from a single plant at the middle of each transect for gas exchange measurements.

```{r, locations, echo=FALSE,message=FALSE,warning=FALSE}
#| label: tbl-locations
#| tbl-cap: "ʻIlima study sites for this study."
#| 
read_rds("../processed-data/site.rds") |>
  filter(site_code != "nnpl") |>
  mutate(Island = case_when(
    island == "oahu" ~ "Oʻahu",
    island == "hawaii" ~ "Hawaiʻi"
  )) |>
  arrange(site_type, site) |>
  select(Site = site, Island, Habitat = site_type, Latitude = latitude_degree,
         Longitude = longitude_degree, `Elevation (mas)` = elevation_m) |>
  knitr::kable(format = "latex", digits = c(rep(0, 3), 3, 3, 0),
               booktabs = TRUE) |>
  kableExtra::kable_styling(latex_options = c("striped", "no_vertical_lines")) |>
  kableExtra::column_spec(1, width = "4cm") |>
  kableExtra::column_spec(2, width = "1.5cm") |>
  kableExtra::column_spec(3, width = "1.5cm") |>
  kableExtra::column_spec(4, width = "1.5cm") |>
  kableExtra::column_spec(5, width = "1.5cm") |>
  kableExtra::column_spec(6, width = "1.5cm")

```

## Leaf traits --- {#leaf-traits}

### Stomata --- {#stomata}

We estimated the stomatal density and size on ab- and adaxial leaf surfaces from all leaves. For pubescent leaves (usually coastal), we dried and pressed leaves for $\approx 1$ week [@hill_temperature_2014], carefully scraped trichomes off with a razor blade, and rehydrated the leaf. For glabrous leaves, we used fresh leaves. We applied clear nail polish to both leaf surfaces of fresh or rehydrated leaves in the middle of the lamina away from major veins. After nail polish dried, we mounted impressions on a microscope slide using transparent tape [@mott_amphistomy_1991]. We digitized a portion of each leaf surface impression using a brightfield microscope (Leica DM2000, Wetzlar, Germany). We counted all stomata and divided by the visible leaf area (0.890 mm$^2$) to estimate density and measured guard cell length from five randomly chosen stomata per field using ImageJ [@schneider_nih_2012].

### Leaf thickness --- {#leaf-thickness}

We cut thin sections using two razor blades taped together and cutting the leaf in a petri dish of water, wet-mounted sections onto a slide, and took digital micrographs using a brightfield microscope, as described above.

### Gas exchange measurements --- {#gas-exchange-measurements}

At each site, we selected one representative leaf from one plant near the middle of the transect for gas exchange measurements using a portable infrared gas analyze LI-6800PF, LI-COR Biosciences, Lincoln, NE, USA). We estimated the photosynthetic rate ($A$) and stomatal conductance to water vapor ($g_\text{sw}$) at saturating light ($\text{photosynthetic photon flux density (PPFD)} = 2000~\mu \text{mol}~\text{m}^{-2}~\text{s}^{-1}$), ambient CO$_2$ (415 ppm), and $T_\mathrm{leaf} = 25.0 \textendash 29.3° \text{C}$. We also estimated 'amphistomy advantage' ($\mathrm{AA}$) *sensu* @parkhurst_adaptive_1978, but with modification. For each leaf, we measured the photosynthetic rate of an untreated amphistomatous leaf ($A_{\mathrm{amphi}}$) over a range of $g_\text{sw}$ values, holding other environmental variables constant. We refer to this as an $A \textendash g_\text{sw}$ curve. We compared the $A \textendash g_\text{sw}$ curve of the untreated leaf to the photosynthetic rate of "pseudohypostomatous" leaf ($A_\mathrm{hypo}$), which is the same leaf but with gas exchange through the upper surface blocked by a neutral density plastic (propafilm). We refer to the untreated and partially blocked leaves as "amphi" and "pseudohypo", respectively. $\mathrm{AA}$ is calculated as the log-response ratio of $A$ compared at the same total $g_\text{sw}$:

$$\mathrm{AA} = \mathrm{log}(A_{\mathrm{amphi}} / A_{\mathrm{hypo}})$$

The log-response ratio is easy to interpret because values above 0 indicate a photosynthetic advantage of amphistomy, whereas values less than 0 indicate a disadvantage. The response ratio is also better than the absolute difference because it indicates a proportional change in $A$, which facilitates comparisons across leaves and environments with different baseline photosynthetic rates. The irradiance of the light source in the pseudohypo leaf is higher because the propafilm reduces transmission. Because the propafilm reduces transmission, we used a higher incident $\mathrm{PPFD}$ for pseudohypo leaves (higher by a factor 1/0.91, the inverse of the measured transmissivity of the propafilm). We also set the stomatal conductance ratio, for purposes of calculating boundary layer conductance, to 0 for pseudohypo leaves following manufacturer directions.

To control for order effects, we alternated between starting with amphi (untreated) or pseudohypo (treated) leaf measurements. In the field, we acclimated the focal leaf to high light ($\mathrm{PPFD} = 2000~\mu \text{mol}~\text{m}^{-2}~\text{s}^{-1}$), high relative humidity ($\mathrm{RH} = 70\%$), $T_\text{leaf} = 25-30~° \text{C}$, depending on ambient temperature, until $A$ and $g_\text{sw}$ reach their maximum. We used these data as our estimates of maximum $A$ and $g_\text{sw}$. After that, we decreased $\mathrm{RH}$ to $\approx 10\%$ to induce rapid stomatal closure without photosynthetic downregulation. We logged data until $g_\text{sw}$ reached its nadir. We then repeated the process with the other treatment (amphi or pseudohypo).

## Data analysis --- {#data-analyis}

### Objective 1: Do coastal leaves have greater amphistomy advantage than montane leaves? ---

It is not feasible to record $A_\mathrm{amphi}$ and $A_\mathrm{hypo}$ at the exact same $g_\text{sw}$. To overcome this, we fit $A \textendash g_\text{sw}$ curves using a flexible generalized additive model (GAM) to interpolate modeled $A$ for amphi and pseudohypo leaves at the same $g_\text{sw}$. In 10 of 12 leaves, the minimum $g_\text{sw}$ of the amphi curve was smaller than the maximum $g_\text{sw}$ of the pseudohypo curve (i.e. the curves overlapped for a range of $g_\text{sw}$ values). In those cases, we estimated $\mathrm{AA}$ at the $g_\mathrm{sw}$ value in the middle of the range of overlap. In 2 of 12 leaves, the $A \textendash g_\text{sw}$ curves did not quite overlap because the minimum $g_\text{sw}$ of the amphi curve was slightly greater than the maximum $g_\text{sw}$ of the pseudohypo curve. In those cases, we estimated $\mathrm{AA}$ by extrapolating slightly, `r gsw_extrap`, beyond the measured curves to the $g_\mathrm{sw}$ value in between the curves. We fit GAMs using the *R* package **brms** version `r fit_licor$version$brms` [@burkner_brms_2017] for Bayesian statistical inference with MCMC sampling in *Stan* [@stan_development_team_stan_2023]. We used CmdStan version `r fit_licor$version$cmdstan` and **cmdstanr** version `r fit_licor$version$cmdstanr` [@gabry_cmdstanr_2023] to interface with *R* version `r getRversion()` [@r_core_team_r:_2023]. We sampled from posterior distribution from `r dim_fit_licor[2]` chains with `r dim_fit_licor[1]` iterations each after `r dim_fit_licor[1]` warmup iterations. We estimated parameters and confidence intervals as the median and 95% quantile intervals of the posterior, respectively. The key prediction is that $\mathrm{AA}_\text{coastal} > \mathrm{AA}_\text{montane}$.

### Objective 2a: Are coastal leaves thicker than montane leaves? ---

We tested whether leaf thickness (log-transformed) varied between coastal and montane populations and among individuals within populations using a Bayesian mixed-effects model with site-type as a fixed effect and individual plant and site as random effects. We used the *R* package **brms** version `r fit_thickness$version$brms` [@burkner_brms_2017] to fit the model in *Stan* [@stan_development_team_stan_2023] with CmdStan version `r fit_thickness$version$cmdstan` and **cmdstanr** version `r fit_thickness$version$cmdstanr` [@gabry_cmdstanr_2023]. We sampled the posterior distribution from `r dim_fit_thickness[2]` chains with `r dim_fit_thickness[1]` iterations each after `r dim_fit_thickness[1]` warmup iterations. We estimated the relationship between population average leaf thickness and $\mathrm{AA}$ measured from a single individual per population. We used this approach most of the variation in leaf thickness was among sites (see Results) and the plant selected for gas exchange measurements was not always among the plants randomly selected for leaf thickness, precluding individual level correlation. We propagated uncertainty about in $\mathrm{AA}$ and leaf thickness by integrating over the entire posterior distribution sample for each variable. The key prediction is that the effect of leaf thickness on $\mathrm{AA}$ is positive.

### Objective 2b: Is $g_{\mathrm{smax,ratio}}$ closer to 0.5 in coastal leaves than montane leaves? ---

We tested whether $g_\mathrm{smax,ratio}$ varied between coastal and montane populations and among individuals within populations using a Bayesian mutliresponse, mixed-effects model. The modeled response variables are stomatal count and guard cell length on each surface. Counts were modeled as negative binomially distributed variable from a latent stomatal density and a parameter $\phi$ to estimate overdispersion in counts relative to a Poisson model. For all traits, the explanatory variables were site-type as a fixed effect and leaf within individual plant, individual plant, and site as random effects. We used the *R* package **brms** version `r fit_stomata$version$brms` [@burkner_brms_2017] to fit the model in *Stan* [@stan_development_team_stan_2023] with CmdStan version `r fit_stomata$version$cmdstan` and **cmdstanr** version `r fit_stomata$version$cmdstanr` [@gabry_cmdstanr_2023]. We interpolated missing adaxial guard cell lengths from `r length(which(fit_stomata$data$upper_number_of_stomata == 0))` out of `r nrow(fit_stomata$data)` samples with zero adaxial stomata using the "mi" function in **brms** package. We sampled the posterior distribution from `r dim_fit_stomata[2]` chains with `r dim_fit_stomata[1]` iterations each after `r dim_fit_stomata[1]` warmup iterations. From each posterior sample, we calculated $g_{\mathrm{smax,ratio}}$ as:

$$g_\mathrm{smax,ratio} = \frac{g_\mathrm{smax,upper}}{g_\mathrm{smax,lower} + g_\mathrm{smax,upper}},$$ where $g_\mathrm{smax,lower}$ and $g_\mathrm{smax,upper}$ are maximum stomatal conductance to water vapor at $T_\mathrm{leaf} = 25 ^ \circ$ C on the lower and upper surface, respectively. The maximum stomatal conductance is calculated from stomatal density and length, assuming that stomata are fully open, following @sack_developmental_2016.

<!-- They defined a biophysical and morphological constant as: -->

<!-- \begin{align*} -->

<!--   b = & D_\text{wv} / v \\ -->

<!--   m = & \frac{\pi c ^ 2}{j ^ {0.5} (4 h j + \pi c)} -->

<!-- \end{align*} -->

<!-- $b$ is the diffusion coefficient of water vapor in air ($D_\text{wv}$) divided by the kinematic viscosity of dry air ($v$). $D_\text{wv} = 2.49 \times 10^{-5} ~ \text{m}^2 ~ \text{s}^{-1}$ and $v = 2.24 \times 10 ^ {-2} ~ \text{m}^3 ~ \text{mol} ^ {-1}$ at $25 ^ \circ$ [@monteith_principles_2013]. For kidney-shaped guard cells, $c = h = j = 0.5$. -->

We estimated the relationship between leaf $g_\mathrm{smax,ratio}$ and $\mathrm{AA}$ measured from a single leaf per population. We propagated uncertainty about in $\mathrm{AA}$ and $g_\mathrm{smax,ratio}$ by integrating over the entire posterior distribution sample for each variable. The key prediction is that the effect of $g_\mathrm{smax,ratio}$ on $\mathrm{AA}$ is positive until $g_\mathrm{smax,ratio} < 0.5$.

# RESULTS --- {#results}

## Amphistomy advantage is greater in coastal leaves ---

Amphistomy increases photosynthesis in leaves of coastal ʻilima plants more than that of montane plants. $\mathrm{AA}$ was significantly greater than 0 (95% confidence intervals did not overlap 0) in 5 of 6 coastal leaves, but only 1 of 6 montane leaves (\autoref{fig:habitat-aa}; see \autoref{fig:licor} for individual curves). Overall, the average $\mathrm{AA}$ among coastal and montane leaves is $`r AA_coastal`$ and $`r AA_montane`$, respectively; the difference in average $\mathrm{AA}$ between habitat types is $\mathrm{AA}_\text{coastal} - \mathrm{AA}_\text{montane} = `r AA_diff`$.  The difference between coastal and montane leaves is lower,  $`r AA_diff1`$, but still greater than 0 if we remove two sites where we extrapolated slightly beyond fitted $A \textendash g_\mathrm{sw}$ curves. Maximum photosynthetic rate was slightly, but not significantly higher in coastal leaves (Welch Two Sample *t*-test, $t_{`r round(fit_habitat_Ags[['A']][['parameter']], 2)`} = `r round(fit_habitat_Ags[['A']][['statistic']], 2)`$, $P = `r round(fit_habitat_Ags[['A']][['p.value']], 2)`$); total stomatal conductance was similar (Welch Two Sample *t*-test, $t_{`r round(fit_habitat_Ags[['gsw']][['parameter']], 2)`} = `r round(fit_habitat_Ags[['gsw']][['statistic']], 2)`$, $P = `r round(fit_habitat_Ags[['gsw']][['p.value']], 2)`$) in coastal and montane leaves (\autoref{fig:habitat-Ags}).

```{=tex}
\begin{figure}[H]
  \includegraphics{../figures/habitat-aa.pdf}
  \caption{Coastal leaves benefit more amphistomy than montane leaves. A positive amphistomy advantage ($y$-axis) means that the photosynthetic rate of an amphistomatous leaf is greater than that of an identical pseudohypostomatous leaf at the same overall $g_\mathrm{sw}$. Each point-interval is the median posterior estimate plus 95\% confidence interval of amphistomy advantage for that leaf. Each leaf is from a different montane (blue) or coastal (orange) site, arranged by habitat and ascending amphistomy advantage within habitat. The longer horizonal bars are the average amphistomy advantage for montane and coastal leaves. $g_\mathrm{sw}$, stomatal conductance to water vapor.}
  \label{fig:habitat-aa}
\end{figure}
```
## Leaf thickness is associated with amphistomy advantage between but not within habitats ---

Coastal ʻilima leaves are $`r thickness_diff`$ $\mu\text{m}$ thicker than their montane counterparts. Although coastal leaves are thicker and have greater $\mathrm{AA}$, there is little relationship between leaf thickness and $\mathrm{AA}$ within habitats (\autoref{fig:traits-aa}A; $\text{slope}~= `r thickness_aa`$).

## $g_\mathrm{smax,ratio}$ is not associated with amphistomy advantage ---

Coastal and montane leaves have similar average $g_\mathrm{smax,ratio}$, the ratio of adaxial (upper) to total $g_\mathrm{smax}$, the anatomical maximum stomatal conductance to water vapor (\autoref{fig:habitat-gmaxratio}); coastal leaves have $`r gmaxratio_sitetype`$ higher $g_\mathrm{smax,ratio}$ than montane leaves, but the 95% confidence intervals overlap 0 difference. The $g_\mathrm{smax,ratio}$ is somewhat bimodal among sites. Some sites in both habitats have leaves with $g_\mathrm{smax,ratio} < 0.07$ and others with $g_\mathrm{smax,ratio} > 0.2$ (\autoref{fig:habitat-gmaxratio}). This is particularly noticeable in montane sites where those on the Big Island of Hawaiʻi all have low $g_\mathrm{smax,ratio}$ whereas those on Oʻahu have relatively high $g_\mathrm{smax,ratio}$. There is no relationship between $g_\mathrm{smax,ratio}$ and $\mathrm{AA}$ in either habitat (\autoref{fig:traits-aa}B; $\text{slope}~= `r gmaxratio_aa`$) in our sample.

```{=tex}
\begin{figure}[H]
  \includegraphics{../figures/traits-aa.pdf}
  \caption{Relationships between leaf amphistomy advantage, (A) $g_\mathrm{smax,ratio}$ and (B) leaf thickness among ʻilima (\textit{Sida fallax}) plants from montane (blue) and coastal (orange) habitats in Hawaiʻi., and amphistomy advantage. A positive amphistomy advantage ($y$-axis) means that the photosynthetic rate of an amphistomatous leaf is greater than that of an identical pseudohypostomatous leaf at the same overall $g_\mathrm{sw}$. Each point-interval is the median posterior estimate plus 95\% confidence interval of the trait value. Each leaf is from a different montane (blue) or coastal (orange) site. Lines are the estimated linear regression of (A) log(leaf thickness) and (B) $g_\mathrm{smax,ratio}$ on amphistomy advantage; ribbons are the 95\% confident bands of the regression. $g_\mathrm{smax,ratio}$, anatomical maximum stomatal conductance ratio; $g_\mathrm{sw}$, stomatal conductance to water vapor.}
  \label{fig:traits-aa}
\end{figure}
```

# DISCUSSION --- {#discussion}

Amphistomy is a seemingly simple way that leaves can increase carbon gain without significant additional water loss, yet it is rare in nature and we do not know why. The strong association between amphiostomy and sunny, open habitats suggests that amphistomy may benefit sun leaves more than shade leaves, but progress has been limited by the lack of evidence that amphiostomy *per se* affects improves photosynthesis in a given leaf. By experimentally blocking gas exchange through the upper surface in a controlled environment, we directly compared an amphistomatous leaf to an otherwise identical pseudohypostomatous leaf. This allows us to quantity the ‘amphistomy advantage’ ($\mathrm{AA}$) holding all else constant. Taking advantage of the of the steep climatic gradients in the Hawaiian archipegalo, we applied this new method to show for the first time that sun leaves benefit more from amphistomy than shade leaves in ʻilima (*Sida fallax*). While this result has not yet been validated in other speciesthere are limits to how much we can generalize to all land plants, our results indicate that part of the reason amphistomatous leaves are found most commonly in high light habitats is that the adaptive benefit is greater in this environment.

If $\mathrm{AA}$ is typically greater in sun leaves than shade leaves, it could partially explain the distribution of amphi- and hypostoamtous leaves, but the precise mechanism(s) require further study. One hypothesis is that the internal airspace conductance, $g_\mathrm{ias}$, from stomata to mesophyll cell walls is lower in thicker sun leaves [@parkhurst_adaptive_1978]. All else being equal, a leaf with lower $g_\mathrm{ias}$ will benefit more from amphistomy. Our results partially support this hypothesis. Coastal ʻilima leaves with high $\mathrm{AA}$ (\autoref{fig:habitat-aa}) are thicker than montane leaves, but the relationship between $\mathrm{AA}$ and leaf thickness within habitats is actually slightly negative (\autoref{fig:traits-aa}A), opposite our prediction. Since coastal and montane leaves differ in many respects besides thickness, we do not have enough data to conclude that leaf thickness explains the variation in $\mathrm{AA}$ between habitats. Alternatively, other biochemical or anatomical differences between coastal and montane leaves may explain why $\mathrm{AA}$ is greater in coastal leaves. The negative relationship between leaf thickness and $\mathrm{AA}$ could be explained if thicker leaves compensated by having a more porous mesophyll and/or less tortuous airspaces [@theroux-rancourt_maximum_2021].

A second natural hypothesis is that amphistomatous leaves with few adaxial (upper) stomata benefit less than those with similar densities on both surfaces. We predicted that leaves with $g_\mathrm{smax,ratio}$ closer to 0.5 would have higher $\mathrm{AA}$ based on biophysical models [@gutschick_photosynthesis_1984]. The logic is that a small number of stomata on the upper surface are insufficient to supply the entire upper mesophyll due to limited lateral diffusion [@morison_lateral_2005]. Our results do not support this hypothesis. Montane leaves from Big Island sites had low $g_\mathrm{smax,ratio}$ and low $\mathrm{AA}$ whereas low montane leaves on Oʻahu had high $g_\mathrm{smax,ratio}$, but similarly low $\mathrm{AA}$ (\autoref{fig:traits-AA}B). Among coastal sites, the site with the lowest $g_\mathrm{smax,ratio}$ had the lowest $\mathrm{AA}$, but there was little variation in $g_\mathrm{smax,ratio}$ among coastal leaves in our sample. We therefore cannot rule out that a larger sample of coastal leaves with greater variance in $g_\mathrm{smax,ratio}$ might support this hypothesis.

Two major implications from our study are that 1) photosynthesis in hypostomatous leaves is likely limited by CO$_2$ concentration drawdown within leaf airspaces; and 2) amphistomy *per se* contributes to, but is not wholly responsible for, higher photosynthetic rates among amphistomatous leaves. The amphistomy advantage we observe in ʻilima leaves implies decreased CO$_2$ supply in pseudohypostomatous leaves because of concentration drawdowns in the leaf airspace. Limited diffusion through the airspace has long been hypothesized to depress photosynthesis in hypostomatous leaves [@parkhurst_diffusion_1994], with empirical support from helox studies [@parkhurst_intercellular_1990]. However, these studies relied on interspecific comparisons of amphi- and hypostomatous leaves that differ systematically in many traits that affect gas exchange and photosynthesis [@xiong_one_2020]. Our experimental approach overcomes this limitation and implies that the drop in CO$_2$ concentration from substomatal cavities to the upper surface depresses photosynthesis.

Among land plants grown in a common garden, amphistomatous leaves have on average nearly $2\times$ higher area-based photosynthetic rates [@xiong_one_2020], naively implying an $\mathrm{AA} \approx \text{log}~2 = 0.69$. This is much higher than our estimate of `r round(habitat_aa$coastal, 2)` among coastal ʻilima leaves. The most likely explanation is that amphistomy is not the only cause of high photosynthetic rate. Indeed, species adapted to open, high light habitats with amphistomatous leaves also have higher concentrations of Rubisco, overall stomatal conductance, and photosynthetic capacity [@smith_leaf_1997; @xiong_one_2020]. For a leaf with high photosynthetic capacity that is well illuminated and hydrated, the major limitation becomes CO$_2$. Under these conditions, amphistomy may substantially increase photosynthesis, as we observe in coastal ʻilima leaves. Selection on increased photosynthesis under similar conditions may explain why crops leaves tend to increase stomatal density ratio during domestication [@milla_shifts_2013].

Two limitations of this study are the small sample sample sizes and experimental design that precludes distinguishing genetic from environmental differences in leaf traits. Understanding the mechanistic basis of higher $\mathrm{A}A$ in sun leaves would require much larger sample sizes. Sun leaves tend to be thicker, more densely packed with mesophyll cells, and have greater photosynthetic capacity and higher stomatal conductance, among other traits [@lambers_plant_2008]. Each of these factors and others potentially modulate $\mathrm{AA}$. Quantifying the contribution of all these factors requires larger samples and additional measurements that are beyond the scope of this study, but exciting avenues for future research on leaf structure-function relations. Although many morphological traits that distinguish coastal and montane ʻilima populations persist in a common environment [@yorkston_interfertility_2006], we cannot distinguish between genetic effects and plastic responses to habitat as causes of difference in $\mathrm{AA}$ because we measured naturally occurring plants *in situ*. While disentangling genetic and plastic contributions is not necessarily important for understanding the distribution of amphistomatous leaves, it would be useful understand genetic and environmental contributions to trait variation. A reciprocal transplant would be able to determine the genetic and environmental contributions, as well their interaction, to trait variance in nature.

# CONCLUSIONS --- {#conclusions}

This study reports the first direct experimental evidence that having stomata open on both leaf surfaces, amphistomy, increases photosynthesis for a given total stomatal conductance, particularly in leaves from the type of open, sunny habitats where this trait is most common. By developing a straightforward gas exchange method to block gas exchange through the upper surface, we directly compared the photosynthetic rate of a leaf with gas exchange through both surfaces or just one, holding all other factors constant. In doing so, we found that coastal leaves of the indigenous Hawaiian ʻilima (*Sida fallax*) enjoyed a greater photosynthetic benefit from amphistomy than nearby montane leaves living in more closed forest. We cannot yet ascribe the difference in amphistomy advantage between coastal and montane leaves to particular physiological or anatomical variation, but this is a promising area for future research.

# Acknowledgments --- {#acknowledgments}

<!-- add tom buckley as coauthor or acknowledge help developing method -->

The authors thank Kasey Barton for feedback on an earlier version of this manuscript, the University of Hawaiʻi honors council for guidance to GT, Tawn Speetjens for access to one montane site, startup funds from the University of Hawaiʻi, NSF Award 1929167 to CDM, and NSF Award 2307341 to TNB.

# Author Contributions --- {#author-ontributions}

GT and CDM contributed equally to all stages of this project; TNB contributed to development of the method and helped edit the manuscript.

# Data Availability Statement ---

Custom scripts are available on a GitHub repository (https://github.com/cdmuir/stomata-ilima) and will be archived on Zenodo with a DOI and stable URL upon publication. Raw data will be deposited on Dryad with a DOI and stable URL upon publication.

# Supporting Information ---

Additional supporting information may be found online in the Supporting Information section at the end of the article

* Appendix S1: Supplemental figures and table

<!-- Placeholders for supplemental figures -->

```{=tex}
\renewcommand\thefigure{S\arabic{figure}}    
\renewcommand\thetable{S\arabic{table}}    
\renewcommand\theequation{S\arabic{equation}}    
\setcounter{figure}{0}    
\setcounter{table}{0}    
\setcounter{equation}{0}
```
```{=tex}
\begin{figure}
  \captionsetup{labelformat=empty}
  \caption{}
  \label{fig:licor}
\end{figure}
```
```{=tex}
\begin{figure}
  \captionsetup{labelformat=empty}
  \caption{}
  \label{fig:habitat-Ags}
\end{figure}
```
```{=tex}
\begin{figure}
  \captionsetup{labelformat=empty}
  \caption{}
  \label{fig:habitat-gmaxratio}
\end{figure}
```
```{=tex}
\begin{table}
  \captionsetup{labelformat=empty}
  \caption{}
  \label{tab:sum-data}
\end{table}
```

# Literature Cited --- {#literature-cited}