### 2023-12-13 ms/si.qmd
preparing final submission

---
title: "Appendix S1: Supplemental figures and table"
format:
  pdf:
    keep-tex: true  
    include-in-header: 
      - preamble.tex
    toc: false
header-includes: |
  \usepackage{fancyhdr}
  \pagestyle{fancy}
  \fancyhead[C]{Triplett \textit{et al.} — \textit{American Journal of Botany} 2024 – Appendix S1}
  \addtokomafont{disposition}{\rmfamily}
  \RedeclareSectionCommand[
    font=\normalfont\Large]{section}
  \RedeclareSectionCommand[
    font=\normalfont\normalsize\bfseries]{subsection}
  \RedeclareSectionCommand[
    font=\normalfont\normalsize\itshape]{subsubsection}
  \RedeclareSectionCommand[
    font=\normalfont\normalsize]{paragraph}
mainfont: Times New Roman
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}

library(readr)

# From 15_archive-data.R
sum_data = read_rds("../objects/sum_data.csv")

```

```{=tex}
\renewcommand\thefigure{S\arabic{figure}}    
\renewcommand\thetable{S\arabic{table}}    
\renewcommand\theequation{S\arabic{equation}}    
\setcounter{figure}{0}    
\setcounter{table}{0}    
\setcounter{equation}{0}
```

```{r, sum-data, echo=FALSE,message=FALSE,warning=FALSE,results='asis'}
#| label: sum-data
#| tbl-cap: "Average traits values among ʻilima leaves at each site. $\\mathrm{SD}_\\text{abaxial}$ is the stomatal density per mm$^2$ on the abaxial (lower) surface; $\\mathrm{SD}_\\text{adaxial}$ is the stomatal density per mm$^2$ on the adaxial (upper) surface; $\\mathrm{GCL}_\\text{abaxial}$ is the guard cell length in $\\mu$m on the abaxial (lower) surface; $\\mathrm{GCL}_\\text{adaxial}$ is the guard cell length in $\\mu$m on the adaxial (upper) surface; Leaf thickness is the length from upper cuticle to lower cuticle in $\\mu$m; $A$ is the photosynthetic rate in $\\mu \\text{mol CO}_2~\\text{m}^{-2}~\\text{s}^{-1}$; $g_\\mathrm{sw}$ is the stomatal conductance to water vapor in $\\text{mol}~\\text{m}^{-2}~\\text{s}^{-1}$."
#| 
sum_data |>
  knitr::kable(format = "latex", digits = c(rep(0, 3), rep(2, 6), 3),
               booktabs = TRUE, escape = FALSE, longtable = TRUE) |>
  kableExtra::kable_styling(latex_options = c("striped", "no_vertical_lines")) |>
  kableExtra::column_spec(1, width = "4cm") |>
  kableExtra::column_spec(2, width = "1.5cm") |>
  kableExtra::column_spec(3, width = "1.5cm") |>
  kableExtra::landscape()

```

\newpage

```{=tex}
\begin{figure}
  \includegraphics{../figures/ags-curve.pdf}
  \caption{Idealized method for collecting $A \textendash g_\text{sw}$ curves on either amphi or pseudohypo leaves. A) After clamping the leaf into the LI-6800 chamber, it acclimates to high light ($\mathrm{PPFD} = 2000~\mu \text{mol}~\text{m}^{-2}~\text{s}^{-1}$) and humidity ($\mathrm{RH} = 70\%$). This induces stomata to open, increasing $g_\mathrm{sw}$ and $A$ until they reach a maximum. We abruptly lower the chamber humidity to $\approx 10\%$ to close stomata and log data (black points) until $g_\mathrm{sw}$ and $A$ reach their nadir. B) We fit $A \textendash g_\text{sw}$ curves to logged data points. The asterisk in both panels indicates the data point used for maximum $A$ and $g_\mathrm{sw}$. C) $\mathrm{AA}$ is low (left panel) when the photosynthetic rate of an amphi leaf is similar to a pseudohypo leaf at the same total $g_\mathrm{sw}$ ($x$-axes); large $\mathrm{AA}$ (right panel) is indicated when an amphi leaf has a higher photosynthetic rate than a pseudohypo leaf. Abbreviations: $A$ is the photosynthetic rate; $\mathrm{AA}$ is the amphistomy advantage; $g_\mathrm{sw}$ is the stomatal conductance to water vapor; $\mathrm{PPFD}$ is photosynthetic photon flux density; $\mathrm{RH}$ is relative humidity.}
  \label{fig:ags-curve}
\end{figure}
```

\newpage

```{=tex}
\begin{figure}
  \includegraphics{../figures/licor.pdf}
  \caption{See next page.}
  \label{fig:licor}
\end{figure}
```

\newpage

\setcounter{figure}{\numexpr\value{figure}-1\relax}

```{=tex}
\begin{figure}
  \caption{(Continued from previous page.) Individual $A \textendash g_\mathrm{sw}$ curves used to estimate $\mathrm{AA}$. For each leaf, one per site, we measured $A$ ($y$-axis) over a range of $g_\mathrm{sw}$ ($x$-axis) on the same leaf with two treatments: `amphi' (circles, solid line) leaves were untreated; `pseudohypo' (squares, dashed line) leaves had no conductance through the upper (adaxial) surface. In all coastal (orange) and montane (blue) leaves, we fit generalized additive models and 95\% confidence ribbons to estimate $\mathrm{AA}$ at a $g_\mathrm{sw}$ where the curves overlap (vertical black line). In leaves from Kaloko-Honokōhau national historical park and Puakō petroglyph park, we extrapolated sligtly beyond fitted curves because they did not quite overlap. Symbols: $\mathrm{AA}$, amphistomy advantage; $A$, photosynthetic rate in $\mu \text{mol CO}_2~\text{m}^{-2}~\text{s}^{-1}$; $g_\mathrm{sw}$, stomatal conductance to water vapor in $\text{mol}~\text{m}^{-2}~\text{s}^{-1}$.}
  \label{fig:licor}
\end{figure}
```

\newpage

```{=tex}
\begin{figure}
  \includegraphics{../figures/pp-licor.pdf}
  \caption{Posterior predictions (thin grey lines) from fitted $A \textendash g_\mathrm{sw}$ curves closely match the observed distribution (thick black line), indicating that the statistical model adequately captures variation in the response variable over the measured range. Symbols: $A$, photosynthetic rate in $\mu \text{mol CO}_2~\text{m}^{-2}~\text{s}^{-1}$; $g_\mathrm{sw}$, stomatal conductance to water vapor in $\text{mol}~\text{m}^{-2}~\text{s}^{-1}$.}
  \label{fig:pp-licor}
\end{figure}
```

\newpage

```{=tex}
\begin{figure}
  \includegraphics{../figures/habitat-Ags.pdf}
  \caption{The photsynthetic rate (left facet) and stomatal conductance to water vapor (right facet) of montane (blue) and coastal (orange) ʻilima leaves. Each point-interval is the median posterior estimate plus 95\% confidence interval of trait value for that habitat. Smaller points next to each point-interval are the $g_\mathrm{smax,ratio}$ of individual plants, one per site. Symbols: $A$, photosynthetic rate in $\mu \text{mol CO}_2~\text{m}^{-2}~\text{s}^{-1}$; $g_\mathrm{sw}$, stomatal conductance to water vapor in $\text{mol}~\text{m}^{-2}~\text{s}^{-1}$.}
  \label{fig:habitat-Ags}
\end{figure}
```

\newpage

```{=tex}
\begin{figure}
  \includegraphics{../figures/habitat-gmaxratio.pdf}
  \caption{The $g_\mathrm{smax,ratio}$ ($y$-axis) of montane (blue) and coastal (orange) ʻilima leaves. Each point-interval is the median posterior estimate plus 95\% confidence interval of $g_\mathrm{smax,ratio}$ for that site. Sites are arranged by habitat and ascending $g_\mathrm{smax,ratio}$ within habitat. Smaller, transparent points next to each point-interval are the $g_\mathrm{smax,ratio}$ of individual plants. Symbols: $g_\mathrm{smax,ratio}$, the ratio of anatomical maximum stomatal conductance to water vapor on the the adaxial surface over the total.}
  \label{fig:habitat-gmaxratio}
\end{figure}
```